Machine learning topological phases in real space

### 2020-06-24 8_prb_submission/1_reply_referee/reply_referee.tex

\newcommand{\genquote}[1]{\refstepcounter{quoter} \label{#1} \ref{#1}}
\newcommand{\citequote}[1]{\ref{#1}}

\definecolor{antiquefuchsia}{rgb}{0.57, 0.36, 0.51}

%\stepcounter{quote}
%\refstepcounter{quote}

\begin{document}
\section*{Reply to the first referee }
\section*{Reply to the second referee}
Dear referee,

we appreciate the valuable insights on how to improve our article that you have provided us with. Below we address each of the emphasized points separately, as they were written in the reply that was sent back to us. We quote the issues raised in italic, and highlight in boldface the points that we thought were particularly relevant.

\subsection*{Major}

\emph{I am not sure about the novelty and generality of their method. For
example, \textbf{their pipeline cannot work for complex experimental physical
systems, where the Hamiltonian could be unknown}\genquote{pipeline_not_work}. Moreover, the authors
state that it is sufficient to use the input data in real space to
predict the topological phase with high accuracy. \textbf{It is not clear to
me what "sufficient" and "high accuracy" mean}\genquote{sufficient_accuracy}, and the advantages of
using the input data in real space instead of the Hamiltonian in
wavevector space. \textbf{I expect the evidence to indicate that their method
is superior or comparable to any other method}\genquote{superior_comparable}. It is worth mentioning
that the author already discussed \textbf{the motivations for developing a
data-driven approach based on real space}\genquote{motif_real_space} in the "Learning topological
phases from real space data" section. However, these arguments should
be pointed carefully in the "Introduction" section.}

@@

\vspace{0.25cm}
\citequote{pipeline_not_work} Mention the possibility of using incomplete information with eigenvector ensembling.

\citequote{sufficient_accuracy} ...

\citequote{superior_comparable} Mention trends in model explainability as well as performance scores.

\citequote{motif_real_space} We have therefore moved the first paragraph of the section "Learning topological phases from real space data" to the introduction.

\vspace{0.5cm}
\emph{I am not sure which crucial problems their algorithm can solve that
was not possible before with machine learning. Furthermore, along with
the emerging research of unsupervised learning methods in realizing
the topological phases, \textbf{why should the supervised learning method be
focused in this context?}\genquote{unsupervised_learning} In the present stage of this manuscript,
where I do not see the proposal's advantages, I would prefer the
\textbf{unsupervised approaches that could grasp information from a given
system without knowing their phases}\genquote{unsupervised_approaches}. In my opinion, a proper intrinsic
route for understanding physical systems without much information on
the dynamics of the system will lead to significant future
developments and a better understanding of physic systems.}

@@

\vspace{0.25cm}
\citequote{unsupervised_learning} Mention semi-supervised learning, model explainability.

\citequote{unsupervised_approaches} Mention Hamiltonian compression, of unsupervised learning through dimensional reduction.

\vspace{0.5cm}
\emph{The details of \textbf{the eigenvector ensembling algorithm should be
addressed in the main text}\genquote{eigenvector_ensembling_alg} instead of supplemental material. What is
the specific algorithm used in the paper to train on eigenvectors? I
think it is important even if the readers are not familiar with some
ML techniques. Some readers in physics may find it difficult to
understand some ML technical terms such as bootstrapping, training and
validation, test sets, etc. In addition, \textbf{I expect proof of what
"physics" their method captures and why the eigenvector can play an
important role in characterizing the topological phase}\genquote{eigenvector_physics}. If this
physical interpretability is not mentioned, it is very difficult to
see the contribution of the method in physics.}

@@

\vspace{0.25cm}
\citequote{eigenvector_ensembling_alg} We have moved the section "The eigenvector ensembling algorithm" to the main paper. The issue of specifying the particular learning algorithms used was also addressed.

\citequote{eigenvector_physics} One of the strongest points of the paper is to introduce model explainability tools from machine learning in physics. The need to use eigenvectors is that they encode all the physical information from a system. The paper paves the way for new theoretical investigation on the relations between the shannon entropy of eigenvectors and topological phase transitions.

\vspace{0.5cm}
\emph{The authors made efforts to analyze how the algorithm was able to
recover the Hamiltonians' global property in the "Information Entropy
Signatures" section. The authors state that learning topological
phases from local real-space data in bulk is still possible even for
small subsets of lattice sites, then refer us to the section "Learning
topological phases from real space data" in the Supplementary
Material. The authors mention on page 4 of the Supplementary Material
that \textbf{"key topological information can be said to be localized on a few
lattice sites," which is a particularly interesting statement to me}\genquote{key_top}.
However, I fail to understand the physical insights behind it. Without
the proper explanation, \textbf{it is difficult to see the effectiveness of
their method or verify the method with a more complex model}\genquote{more_complex_model} instead of
the simple SSH form.}

@@

\vspace{0.25cm}
\citequote{key_top} Move this sentence to main paper.

\citequote{more_complex_model} Add FFT of entropy signatures? Mention entropic uncertainty? Talk about information gain? Mention that interpretability in ML is often challenging and a fast evolving field. Furthermore, eigenvector ensembling can always be used as a preprocessing step to other ML models, such as in semi supervised learning tasks.

\subsection*{Minor}

\emph{The authors state in the abstract that "model explainability in
machine learning can advance the research of exotic quantum materials
with properties that may power future technological applications such
as qubit engineering for quantum computing." \textbf{Could you explain a bit
more about properties that may power future quantum computing?}\genquote{quantum_comp}}

\vspace{0.25cm}
\citequote{quantum_comp} Insert comments about topological qubits?

\vspace{0.5cm}
\emph{On page 4, the authors state that "Figures 2 and 3 illustrate single
iterations of experiments 1 and 2 as seen from parameter space". \textbf{What
are the experiments 1 and 2?}\genquote{experiments_1_2}}

\vspace{0.25cm}
\citequote{experiments_1_2} Experiments 1 and 2 were mentioned as subtitles to the sections.

\vspace{0.5cm}
\emph{\textbf{The authors should explain the evaluation metric in the main text}\genquote{evaluation_metrics},
such as "the accuracy" and "the probability heatmap," "eigenvector
accuracy," and "Hamiltonian accuracy."}

\vspace{0.25cm}
\citequote{evaluation_metrics}

\vspace{0.5cm}
\emph{I prefer to \textbf{put the definition of information entropy signatures in
the main}\genquote{inf_entropy_sig} text to help the readers understand what the method tries to do.}

\vspace{0.25cm}
\citequote{inf_entropy_sig}

\end{document}

### 2020-06-24 8_prb_submission/1_reply_referee/article_with_comments/article.tex
Added comments about sampling and dimensionality reduction

\emph{I am not sure about the novelty and generality of their method. For
example, \textbf{their pipeline cannot work for complex experimental physical
systems, where the Hamiltonian could be unknown}\genquote{pipeline_not_work}. Moreover, the authors
state that it is sufficient to use the input data in real space to
predict the topological phase with high accuracy. \textbf{It is not clear to
me what "sufficient" and "high accuracy" mean}\genquote{sufficient_accuracy}, and the advantages of
using the input data in real space instead of the Hamiltonian in
wavevector space. \textbf{I expect the evidence to indicate that their method
is superior or comparable to any other method}\genquote{superior_comparable}. It is worth mentioning
that the author already discussed \textbf{the motivations for developing a
data-driven approach based on real space}\genquote{motif_real_space} in the "Learning topological
phases from real space data" section. However, these arguments should
be pointed carefully in the "Introduction" section.}

\vspace{0.25cm}
\citequote{pipeline_not_work} Mention the possibility of using incomplete information with eigenvector ensembling.

@@

\vspace{0.25cm}
\textcolor{blue}{\citequote{unsupervised_learning}, \citequote{unsupervised_approaches} Our main goal in the article is to investigate the localizability of topological information along lattices of SSH systems via model explainability tools in machine learning. The most straightforward way to do this is by performing supervised learning with decision tree-based algorithms, since their use of Shannon entropy as loss function provides a clear way to measure feature importance. This is compounded by the fact that Shannon entropy plays a very important role in information theory and physics, as is evidenced by the entropic uncertainty relations included in this new version of the article. It should be noted, however, that the eigenvector ensembling algorithm can equally be combined with unsupervised learning algorithms as well. This point is now explicitly mentioned in this new version of the paper, in the last paragraph of the section "THE EIGENVECTOR ENSEMBLING ALGORITHM" which was moved to the main body of the paper.}

@@

\vspace{0.25cm}
\textcolor{blue}{\citequote{eigenvector_ensembling_alg} We have moved the section "The eigenvector ensembling algorithm" to the main paper. The issue of specifying the particular learning algorithms used was also addressed in the same section.}

\citequote{eigenvector_physics} One of the strongest points of the paper is to introduce model explainability tools from machine learning in physics. The need to use eigenvectors is that they encode all the physical information from a system. The paper paves the way for new theoretical investigation on the relations between the shannon entropy of eigenvectors and topological phase transitions.

@@

\vspace{0.25cm}
\citequote{key_top} Move this sentence to main paper.

@@

These results demonstrate that learning topological phases from local real space data in the bulk is still possible even for small subsets of lattice sites. We refer the reader to the section \textbf{Learning topological phases from real space data} in the Supplementary Material for a discussion of how this is possible.

@@

The eigenvector ensembling algorithm employed in this work is likely to have further applications in data-driven physics. This is because most of physics is based on eigenvector decomposition, and statistical physics itself can be seen as an application of similar ensembling principles. It is therefore expected that a much broader class of data-driven physics problems could benefit from the techniques described in this paper.

>>>>

\emph{\textbf{I am not sure about the novelty and generality of their method.}\genquote{pipeline_not_work} For
example, their pipeline cannot work for complex experimental physical
systems, where the Hamiltonian could be unknown. Moreover, the authors
state that it is sufficient to use the input data in real space to
predict the topological phase with high accuracy. \textbf{It is not clear to
me what "sufficient" and "high accuracy" mean}\genquote{sufficient_accuracy}, and the advantages of
using the input data in real space instead of the Hamiltonian in
wavevector space. \textbf{I expect the evidence to indicate that their method
is superior or comparable to any other method}\genquote{superior_comparable}. It is worth mentioning
that the author already discussed \textbf{the motivations for developing a
data-driven approach based on real space}\genquote{motif_real_space} in the "Learning topological
phases from real space data" section. However, these arguments should
be pointed carefully in the "Introduction" section.}

\vspace{0.25cm}
\textcolor{blue}{\citequote{pipeline_not_work} We can think of several use cases for the eigenvector ensembling procedure outlined in the paper. Among these, we emphasize its applicability in tackling the so-called "curse of dimensionality" that plagues many problems in condensed matter and statistical physics. In the paper, we have demonstrated an application in which the dimensionality of an SSH lattice can be greatly reduced in so far as its topological properties are concerned, simply by extracting the most relevant lattice sites in its information entropy signature. Similar feature extraction and dimensionality reduction procedures are often employed in machine learning applications and are surely to become the norm in data-driven physics. The eigenvector ensembling algorithm can thus be viewed as a general framework in this direction. Similarly, by thoughtfully attaching a probability distribution to the eigenvectors, we end up with a powerful sampling technique that may ultimately lead to great reduction in the dimensionality of a Monte Carlo simulation, while still capturing essentially all the relevant physics pertaining to the problem. We have added a paragraph to the \textbf{Discussion} section outlining these possibilities. As for the novelty of the method, the eigenvector ensembling algorithm together with the explainability powers of decision trees allowed us to rigorously investigate the information entropy of topological phase transitions. To our knowledge, our work pioneers the discovery of new concepts in physics through model explainability in machine learning.  
}

@@

\vspace{0.25cm}
\textcolor{blue}{\citequote{unsupervised_learning}, \citequote{unsupervised_approaches} Our main goal in the article is to investigate the localizability of topological information along lattices of SSH systems via model explainability tools in machine learning. We developed a pipeline that accomplishes this task in a very straightforward way by performing supervised learning with decision tree-based algorithms, since their use of Shannon entropy as loss function provides a clear way to measure feature importance. This is compounded by the fact that Shannon entropy plays a very important role in information theory and physics, as is evidenced by the entropic uncertainty relations included in this new version of the article. It should be noted, however, that the eigenvector ensembling algorithm can equally be combined with unsupervised learning algorithms as well. This point is now explicitly mentioned in this new version of the paper, in the last paragraph of the section "The eigenvector ensembling algorithm" which was moved from the Supplementary Material to the main paper. Finally, as we mentioned in our reply to \citequote{pipeline_not_work}, our procedure can be used for example as a powerful dimensionality reduction tool, and therefore many current }

@@

\vspace{0.25cm}
\textcolor{blue}{\citequote{eigenvector_ensembling_alg} We have moved the section "The eigenvector ensembling algorithm" to the main paper. The issue of specifying the particular learning algorithms used to train on eigenvectors was also addressed in the same section.}

\citequote{eigenvector_physics} One of the main themes of the paper is to use model explainability tools from machine learning in to arrive at new concepts in physics. The need to use eigenvectors is that they encode all the physical information from a system. The paper paves the way for new theoretical investigation on the relations between the shannon entropy of eigenvectors and topological phase transitions.

@@

\vspace{0.25cm}
\textcolor{blue}{\citequote{key_top} We agree that this sentence greatly synthesizes the results presented in the paper about the possibility of localizing topological information in SSH lattices. We have therefore moved it to the final patragraph of the section "Information entropy signatures" in the main text.}

@@

These results demonstrate that learning topological phases from local real space data in the bulk is still possible even for small subsets of lattice sites. \textcolor{blue}{In this sense, key topological information can be said to be localized on few sites in the lattice.}\citequote{key_top} We refer the reader to the section \textbf{Learning topological phases from real space data} in the Supplementary Material for a discussion of how this is possible.

@@

The eigenvector ensembling algorithm employed in this work is likely to have further applications in data-driven physics. This is because most of physics is based on eigenvector decomposition, and statistical physics itself can be seen as an application of similar ensembling principles.

\textcolor{blue}{As a concrete example, the study of several many-body systems of current interest in condensed matter physics is hindered by their large dimensionality. This problem, known as \emph{the curse of dimensionality} in the scientific computing community, arises from the necessity of collecting or processing exponentially larger amounts of data as the feature space dimensionality of a problem grows. An approach based on eigenvector ensembling can be of use in such situations both as a dimensionality reduction tool and as a sampling strategy. The first case was illustrated in this work, where it was shown that relevant topological information of SSH systems can be retrieved from few sites in a lattice, which can be exploited as a dimensionality reduction strategy. The latter case, which was not explored here, also poses interesting possibilities, such as sampling eigenstates according to a desired distribution in Monte Carlo simulations of condensed matter systems. Indeed, sampling eigenvectors from a carefully designed probability distribution can ultimately lead to a great reduction in the dimensionality of a problem while still capturing all the relevant physics of a system.}\citequote{pipeline_not_work}  We therefore expect that a much broader class of data-driven physics problems could benefit from the techniques described in this paper.
